The human in the loop: volunteer computing as a socio-technical system

### 2016-02-28 Alife16/paper.tex
Changing stuff
Minor corrections and two questions related with the metrics to detect the patterns for issue #102

\begin{abstract}
Volunteer computing is a form of distributed computing where users
decide on their participation and the amount of time and other resources they will
``lend''. This makes them an essential part of the algorithm and of
the performance of the whole system. As a sociotechnical system, this
participation follows some patterns and in this paper we examine the
result of several volunteer distributed evolutionary computation
experiments and try to find out which are those patterns and what
makes an experiment successful or not, including the feedback loop
that is created between the users and the algorithm itself.
\end{abstract}

@@

Some time ago, our group faced the problem of diminishing funds for
buying new hardware. This was aggravated by the increasing maintenance
costs and extended downtime resulting from the continuous failures of
existing clusters.  Considering this, we leveraged our experience in
the design of web applications with JavaScript and other volunteer and
unconventional distributed evolutionary computing systems to design
and release a new free framework that would allow anyone to create a
volunteer distributed evolutionary computation (EC) experiment using
cloud resources as servers and browsers as clients. This framework was
called NodIO \citep{2016arXiv160101607M}. NodIO provides server
infraestructure for volunteer-based distributed evolutionary computing
experiments by providing a chromosome {\em pool}. This pool is used by
clients in browsers and any other using the application programming
interface (API) to put chromosomes and retrieve them, working then as
a loose, asynchronous and {\em ad hoc} connection among all clients
using it.

This loose connection provides a low-overhead way to connect desktop
experiments with volunteer-based ones, with all of them contributing
to the pool, but every one of them working as separate island 
carrying out their own evolutionary algorithms. This is why NodIO is
proposed mainly as a {\em complement} to existent resources such as
desktop systems or laptops. As long as it provides a non-null
computational capability that can help existing resources find the
solution faster it will have found its purpose. The main use case is
someone setting up NodIO in the cloud, writing a fitness evaluation
function and running a client from his or her own computer, but
requesting help in social networks for additional resources. That is
why, in this system,  we have considered the whole {\em social} aspect
in the design, with issues related to security, trust and privacy
among others. The computing system becomes a {\em techno-social
  system} \citep{vespignani2009predicting}. In this paper we are going
to focus on measuring the response of users to experiments, but at the
same time we will also focus on the technical aspects of the server and how
those might change the behavior of users, improving the capability of % Mario: How these might change
the system.

>>>>

\begin{abstract}
Volunteer computing is a form of distributed computing where users
decide on their participation and the amount of time and other resources they will
``lend''. This makes them an essential part of the algorithm and of
the performance of the whole system. As a sociotechnical system, this
participation follows some patterns and in this paper we examine the
result of several volunteer distributed evolutionary computation
experiments and try to find out which are those patterns and what
makes an experiment successful or not, including the feedback loop
that is created between the users and the algorithm itself.
%Pablo: The patterns we are looking for are... 
\end{abstract}

@@

Some time ago, our group faced the problem of diminishing funds for
buying new hardware. This was aggravated by the increasing maintenance
costs and extended downtime resulting from the continuous failures of
existing clusters.  Considering this, we leveraged our experience in
the design of web applications with JavaScript and other volunteer and
unconventional distributed evolutionary computing systems to design
and release a new free framework that would allow anyone to create a
volunteer distributed evolutionary computation (EC) experiment using
cloud resources as servers and browsers as clients. This framework was
called NodIO \citep{2016arXiv160101607M}. NodIO provides server
infrastructure for volunteer-based distributed evolutionary computing
experiments by providing a chromosome {\em pool}. This pool is used by
clients in browsers and any other using the application programming
interface (API) to put chromosomes and retrieve them, working then as
a loose, asynchronous and {\em ad hoc} connection among all clients
using it.

This loose connection provides a low-overhead way to connect desktop
experiments with volunteer-based ones, with all of them contributing
to the pool, but every one of them working as separate island 
carrying out their own evolutionary algorithms. This is why NodIO is
proposed mainly as a {\em complement} to existing resources such as
desktop systems or laptops. As long as it provides a non-null
computational capability that can help existing resources find the
solution faster it will have found its purpose. The main use case is
someone setting up NodIO in the cloud, writing a fitness evaluation
function and running a client from his or her own computer, but
requesting help in social networks for additional resources. That is
why, in this system,  we have considered the whole {\em social} aspect
in the design, with issues related to security, trust and privacy
among others. The computing system becomes a {\em techno-social
%Pablo: In abstract is named 'sociotechnical'. Which one is the correct form for the rest of the paper?
  system} \citep{vespignani2009predicting}. In this paper, we are going
to focus on measuring the response of users to experiments, but at the
%Pablo: which metric are we going to measure this response?
same time we will also focus on the technical aspects of the server and how
those might change the behavior of users, improving the capability of % Mario: How these might change
the system.
